---
format:
  html: 
    embed-resources: true
    self-contained-math: true
    toc: true
---

## Price Elasticity

Price Elasticity is a measurement of changing quantity due to changing prices.

We will use $\eta_p$ to mean price elasticity.

-   $|\eta_p| < 1$ is an elastic product\
-   $|\eta_p| \le 1$ is an inelastic product

Price elasticity is defined as

$$\eta_p = \frac{\%\Delta{Q}}{\%\Delta{P}}$$ The midpoiont formula for elasticity is more commonly used, and is:

$$
\%\Delta{Q}=\frac{q_2-q_1}{(q_2+q_1)/2} \\
\%\Delta{Y}=\frac{p_2-p_1}{(p_2+p_1)/2} 
$$

### Calculating by Hand

Give the following table:

```{r}
sales = data.frame(price = c(10, 15), qty = c(120, 100))
sales
```

We would calculate $\eta_p$ as

```{r}
((sales$qty[2] - sales$qty[1]) / ( (sales$qty[2] + sales$qty[1]) / 2 )) /
((sales$price[2] - sales$price[1]) / ( (sales$price[2] + sales$price[1]) / 2 ))
```

As a function:

```{r}
pct_delta <- function(x1, x2) {
  (x2 - x1) / mean(c(x2, x1))
}

elasticity_p <- function(prices, qty) {
   pct_delta(qty[1], qty[2]) / pct_delta(prices[1], prices[2]) 
}

elasticity_p(sales$price, sales$qty)

```

### Using a Regression

$$\eta_p = \beta_{price}(\frac{\bar{P}}{\bar{Q}})$$

Using our above sample:

```{r}
model <- lm(qty ~ price, data = sales)

model$coefficients[["price"]] * ( mean(sales$price) / mean(sales$qty) )

```

### Logarithmic Transformation

$$
log(Y) = a + \beta{x} \\ d(log_y)=\beta dX \\ \frac{dY}{y} = \beta{}dX \\ 100b = \frac{\%\Delta Y}{Unit\ \Delta X}
$$

```{r}
model_log <- lm(log(qty) ~ price, data = sales)

model_log$coefficients[["price"]]
```

